---
title: "Danish Kings Portfolio 2"
author: "Amanda Gulstad Pedersen"
date: "2025-26-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load library and kings streadsheet

The spreadsheet DanishMonarchs.csv is separated by semi colon.  

2. Create a `kings` object in R with the different functions below and inspect the different outputs.
-  `read.csv()`
-  `read_csv()`
-  `read.csv2()` 
-  `read_csv2()`
```{r eval=TRUE}

library(tidyverse)

kings1 <- read.csv("DataKings/DanishMonarchs.csv", na = c("NA"))

kings2 <- read_csv("DataKings/DanishMonarchs.csv", na = c("NA"))

kings3 <- read.csv2("DataKings/DanishMonarchs.csv", na = c("NA"))

kings4 <- read_csv2("DataKings/DanishMonarchs.csv", na = c("NA"))
```
Answer: 
1.
The function using an underscore (_) is from the TidyVerse package, while using full stop (.) is a part of base R. Weather to use 1 or 2 depends on which separator the spreadsheet has. As my spreadsheet is separated by a semicolon, using the function read_csv2 is the best option.

2. The result of running `class()` on the `kings` object created with a tidyverse function (which would be kings4) is: [1] "spec_tbl_df" "tbl_df"      "tbl"         "data.frame"

3. Using the function ncol(), we can se the spreadsheet has 5 columns. If using the wrong function for the seperater, the spreadsheet is read as having only 1.

4. Using the function glimpse() we can see a snippet of the dataset and how R interprets the data.

```{r eval = TRUE}


kings <- kings4

class(kings)


ncol(kings)


glimpse(kings)


```


## Calculate the duration of reign for all the kings using the mutate function

Here, I am using the mutate function to create a new column showing the duration of reign for each monarch. I am using !is.na so it only calculates the reign duration of monarchs where both start and end of reign year is known.


```{r eval=TRUE}


kings <- kings %>% 
  filter(!is.na(end_reign_year), !is.na(start_reign_year)) %>% 
  mutate(duration = end_reign_year - start_reign_year)


```

## Calculate the average duration of reign for all rulers


```{r eval=TRUE}



kings_gennemsnit <- kings %>% 
  summarize(gennemsnit_duration = mean(duration, na.rm=TRUE))


```

## How many and which kings enjoyed a longer-than-average duration of reign?
Using the filter() function to create a new object that shows the kings with a longer reign than average. It shows that 25 kings had a reign longer than the average of 19 years.

```{r eval =TRUE}


kings_over_gennemsnit <- filter(kings, duration > mean(duration, na.rm=TRUE))

count(kings_over_gennemsnit)

```


## How many days did the three longest-ruling monarchs rule? 

To create a new object which shows the top 3 longest ruling monarchs, i first use the arrange function to arrange the kings in descending order, and using the slice function, i remove all but the first 3 monarchs. With the use of mutate, we can add a 7th column that shows the duration in days.


```{r eval=TRUE}


Top_3 <- kings %>% 
  arrange(desc(duration)) %>% 
  slice(1:3) %>% 
  mutate(days = duration * 365.25)


```


## Challenge: Plot the kings' duration of reign through time

Using ggplot we can create visuals of the monarch's duration of reign with two different functions, geom_point and geoom_smooth.

```{r eval=TRUE}


kings_ggplot <- kings %>% 
  mutate(midyear = start_reign_year + (end_reign_year - start_reign_year)/2)

kings_ggplot %>% 
  ggplot(aes(x = midyear, y = duration)) +
  geom_point() +
  labs(title = "Duration of Reign Through Time", x = "Reign Midyear", y = "Duration in Years") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        text = element_text(size = 16),
        plot.title = element_text(size = 20))

kings_ggplot %>% 
  ggplot(aes(x = midyear, y = duration)) +
  geom_smooth() +
  labs(title = "Duration of Reign", x = "Reign Midyear", y = "Duration in Years") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        text = element_text(size = 16),
        plot.title = element_text(size = 20))


```


